Credit Risk Distribution and Capital Savings

### 2020-04-02 Analysis/Draft/Draft.Rmd
Last version of draft before switching to bookdown

---
title: "Credit Risk Distribution and Capital Savings"
author:
- name: Lucas Avezum
  affiliation: CentER and EBC, Tilburg University; l.avezum@uvt.nl; +31134664531;  Warandelaan 2, 5037 AB Tilburg, NL, The Netherlands 
date: 
header-includes:
   - \usepackage{rotating, graphicx, dcolumn, longtable}
output:
  pdf_document:
    citation_package: natbib
    latex_engine: pdflatex
    template: tex-ms.tex
    fig_caption: true
  html_document:
    df_print: paged
  word_document: default
bibliography: ../../Auxiliar/mybibfile.bib
citecolor: black
anonymous: false
fontsize: 11pt
spacing: "double"
geometry: margin= 1in
keywords: Internal ratings-based models, Risk-weighted assets dispersion, Procyclicality
JEL: G21, G28, G11
linkcolor: black
biblio-style: apsr
subtitle: ''
thanks: Preliminary draft.  
abstract: "The Basel II accord introduced a concave mapping from probabilities of default (PD) to capital requirements for banks adopting the internal ratings-based (IRB) approach. Therefore, total capital requirements depends not only on the portfolio's average PD but also on the shape of its PD distribution. Using a hand-collected dataset, I propose a method that decomposes the contribution of the PD distribution to capital requirements into an average component and a shape component. While a larger average component increases capital charges, a larger shape component generates capital savings. I document that these capital savings are important sources of risk-weight variability across banks. I further investigate the relationship of capital savings with the business cycle and find it to be countercyclical. The results suggest that a recalibration of the current IRB framework can significantly reduce the procyclicality of capital requirements."
urlcolor: black
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo      = FALSE, 
                      warning   = FALSE, 
                      error     = FALSE, 
                      tidy      = TRUE, 
                      tidy.opts = list(width.cutoff=50), 
                      collapse  = TRUE,
                      warning   = FALSE,
                      error     = FALSE,
                      message   = FALSE, 
                      comment   = "",
                      cache     = FALSE)
# These options are tuned for manuscript/presentation. 
# They basically run R in the background except for spitting out figures/tables
# They also cache results so that you don't have to run your whole analysis every time you fix a typo

@@

library(tidyverse)            ## Data manipulation, pipe operator
library(data.table)           ## Command to bind lists
library(ggridges)             ## Density ridges plot
library(stargazer)            ## Latex tables
library(lfe)                  ## High-dimensional fixed effects
library(gridExtra)            ## Create grid for plots and tables
library(cowplot)              ## Create grid for plots and tables
library(DescTools)            ## Command to winsorize variables
library(ggpubr)               ## More plots
library(textablr)
library(sandwich)
library(lmtest)
#library(relaimpo)            ## Calculate relative importance measures

source("../../Auxiliar/Function.R") ## Capital requirements functions

@@

@@

\begin{equation}
\label{eq4}
\text{RW}^r=\frac{\text{RW}^c}{\text{RW}}.
\end{equation}

\begin{equation}
\label{eq3}
\text{RWA}^s=\text{RWA}^c-\text{RWA}.
\end{equation}

## Capital Savings and Cyclicality

\begin{table}[!htbp] \centering 
  \caption{Cyclicality of Capital Savings} 
  \label{tab1} 
\small 
\begin{tabular}{@{\extracolsep{3pt}}lD{.}{.}{-3} D{.}{.}{-3} D{.}{.}{-3} D{.}{.}{-3} D{.}{.}{-3}} 
\hline 
\hline \\[-1.8ex] 
\\[-1.8ex] & \multicolumn{3}{c}{States}  & \multicolumn{2}{c}{$\Delta$ Initial}  \\ \cline{2-4}\cline{5-6}
\\[-1.8ex] & \multicolumn{1}{c}{Initial} & \multicolumn{1}{c}{A} & \multicolumn{1}{c}{B} & \multicolumn{1}{c}{A} & \multicolumn{1}{c}{B} \\ 
\hline \\[-1.8ex] 
 $\text{RWA}$  & 63.88 & 94.21 & 90.29 & 30.32 & 26.42 \\ 
  $\text{RWA}^c$ & 64.65 & 94.65 & 94.65 & 29.96 & 29.96  \\ 
  $\text{RWA}^s$ & 0.81 & 0.45 & 4.35 & -0.36 & 3.54  \\ 
  \hline \\[-1.8ex]
 Avg. PD & 0.15\% & 0.25\% & 0.25\% &  &   \\ 
  S.d. PD & 0.05\% & 0.05\% & 0.23\% &  &   \\ 
\hline \\[-1.8ex]

\end{tabular} 
\end{table}

@@

Consider now a credit risk shock. In state A, both safer and riskier assets are affected, having their PDs raised to 0.2% and 0.3%, respectively. In state B, only the riskier asset is affected, having its PD equal to 0.4%. Note that the average PD in both states are equal at 0.25%. However, relative to the initial state, the capital savings measure for state A decreases to \$0.45 while it increases to \$4.35 in state B. The mechanism of this example is simple: relative to the initial state, in state A only the location parameter of the PD distribution changes (standard deviation is equal to 0.05% in both states) while in state B also other moments change (standard deviation increases to 0.15%). Note that in both state A and B total capital requirements increase as average credit risk is higher relative to the initial state in both states. This net procyclical effect is in line with the results of @behn2016procyclical. However, due to the increased portfolio dispersion, capital savings are countercyclical in state B. Conversely, because the PD-RW formula elasticity is lower for higher PD estimates, in stata A *Capital savings* are procyclical. Therefore, capital savings are procyclical in state A. Table \ref{tab1} summarizes the example and Figure Figure \ref{fig:pdmapcases}

@@

\pagebreak

@@

```{r, fig.height = 7.9, fig.width = 6, results="hide",fig.cap= cross.dec_cap1, include=TRUE}
 plot1 <- cross.section.decomposition %>% 
  filter(q.SA<1)%>%
  mutate(total= -PD.gain)%>%
  select(name, Country,total, int.gain, ext.gain) %>%
  gather(condition, value, contains(".gain"), factor_key=TRUE)%>%
  mutate(condition = factor(condition, levels = c("int.gain","ext.gain"),
                                       labels = c("Intensive margin", "Extensive margin")))%>%
ggplot(mapping = aes(x = reorder(name, total),
                       y = value)) +
geom_bar(stat = 'identity')+
  #geom_line(aes(y = -total,group=1))  +
  labs(x=NULL, y= NULL) +
    coord_flip() +
  theme_bw() +
  theme(panel.grid.minor = element_blank())+
  facet_wrap(~condition)

plot2 <- cross.section.decomposition %>% 
  filter(q.SA<1)%>%
  select(name, Country, int.gain, ext.gain) %>%
  mutate_if(is.numeric,Winsorize,probs = c(0.02, 1) ,na.rm = TRUE)%>% 
  ggplot(mapping = aes(x = int.gain,
                       y = ext.gain)) +
    geom_point()+
    geom_smooth(method = "lm", color = "black")  +
  labs(x=NULL, y= NULL) +
   theme_bw() +
  theme(panel.grid.minor = element_blank())+
  scale_fill_grey()+
  stat_cor()

@@

```{r, fig.height = 7.9, fig.width = 6, results="hide",fig.cap= cross.dec_cap2, include=TRUE}
plot1 <- cross.section.decomposition %>% filter(q.SA<1)%>%
  mutate(total= -PD.gain)%>%
  select(name, Country,total, other.gain, PD.gain) %>%
  gather(condition, value, contains(".gain"), factor_key=TRUE)%>%
  mutate(condition = factor(condition, levels = c("PD.gain","other.gain"),
                                       labels = c("Capital savings", "Other components")))%>%
ggplot(mapping = aes(x = reorder(name, total),
                       y = value)) +
geom_bar(stat = 'identity')+
  labs(x=NULL, y= NULL) +
    coord_flip() +theme_bw()+
  theme(panel.grid.minor = element_blank())+
  facet_wrap(~condition)

plot2 <- cross.section.decomposition %>% 
  filter(q.SA<1)%>%
  select(name, Country, other.gain, PD.gain) %>%
  mutate_if(is.numeric,Winsorize,probs = c(0.02, 1) ,na.rm = TRUE)%>% 
  ggplot(mapping = aes(x = PD.gain,
                       y = other.gain)) +
    geom_point()+
    geom_smooth(method = "lm", color = "black")  +
  labs(x=NULL, y= NULL) +
   theme_bw() +
  theme(panel.grid.minor = element_blank())+
  scale_fill_grey()+
  stat_cor(label.y= -0.1)

@@

unrestrict <- cross.section.decomposition %>% filter(q.SA<1) %>%
  mutate(epsilon = rnorm(nrow(.),0,0.0000001),
         total.gain = total.gain+epsilon)%>% 
  lm(data = ., total.gain ~ IRB.gain + mix.gain + rollout.gain + int.gain + ext.gain + SA.gain)%>%
  relaimpo::calc.relimp(type = c("lmg"), rela = FALSE )

@@

The follow-up to the 2008 financial crisis offer an plausible explanation for the dynamics of *Capital savings* in time, which is illustrated in Figure \label{fig:timevariables}. Increased supervisory pressure to increase capital ratios (top plot), especially after the Basell III, led banks to deleverage and reduce risk-taking.\footnote{Although supervisors favored raising additional capital as means to achieve the required capital ratios \citep{EBA2011recommendation}, banks respond to capital requirements mostly by reducing RWA \citep{gropp2018banks}}  The shift for safer assets not only reduced the average PD on banks' portfolios (middle plot) but also compressed their PD distribution (bottom plot). This distribution compression leads to a reduction in *Capital savings*. This explanation is in line with the hypothesis that *Capital savings* are countercyclical. Efforts to increase capital ratio by risk reduction are increasingly more difficult.

@@

\pagebreak

```{r , fig.height = 10, fig.width = 7, results="asis",out.width = "1\\textwidth"}
p_w <- c(0.05, 0.95)         ## Set winsorizing limits
bank.data.win <- bank.data %>% 
  ungroup()%>%
  filter(!is.na(RW.s), bvdid %in% pillar3.data$bvdid) %>% 
  mutate_if(is.numeric,Winsorize,probs = p_w ,na.rm = TRUE)%>% 
  group_by(bvdid)

bank.data.win %>% select(
  log.CAR_gr, 
  log.K_gr, 
  log.RWA.IRB_gr, 
  log.RWA.hat_gr, 
  log.RWA.s_gr,
  log.RWA.r_gr,
  log.RWA.r.Retail_gr,
  log.RWA.r.Wholesale_gr, 
  log.RWA.r.Corporate_gr, 
  log.RWA.r.Sovereign_gr, 
  log.RWA.r.Banks_gr, 
  mean.PD_gr, 
  var.PD_gr, 
  skew.PD_gr,
  kurt.PD_gr, 
  gini.PD_gr,
  log.gdp_gr,
 #log.asset_gr, 
 #capital.ratio_gr, 
 #loans.ratio_gr, 
 #NII.ratio_gr,  
 #deposit.ratio_gr,
 #income.ratio_gr,
  log.gdp,
  log.asset,
  capital.ratio,
  income.ratio,
  loans.ratio, 
  NII.ratio, 
  NPL.ratio,
  deposit.ratio) %>%
  as.data.frame() %>% 
stargazer(header = FALSE, type="latex", single.row = TRUE, column.sep.width = "1pt", font.size = "small",
          label = "tab:summary",
          omit.summary.stat = c("p25","p75"),
          covariate.labels = 
            c("$\\Delta \\text{Log Capital Ratio}$",
              "\\hspace{0.5cm}$\\Delta \\text{Log capital}$",
              "\\hspace{0.5cm}$\\Delta \\text{Log RWA}$",
              "\\hspace{1.0cm}$\\Delta \\text{Log RWA}^c$",
              "\\hspace{1.0cm}$\\Delta \\text{Log RWA}^s$",
              "\\hspace{1.0cm}$\\Delta \\text{Log RWA}^r$",
              "\\hspace{1.5cm}$\\Delta \\text{Log RWA}^r$ Retail",
              "\\hspace{1.5cm}$\\Delta \\text{Log RWA}^r$ Wholesale",
              "\\hspace{2.0cm}$\\Delta \\text{Log RWA}^r$ Corporate",
              "\\hspace{2.0cm}$\\Delta \\text{Log RWA}^r$ Sovereign", 
              "\\hspace{2.0cm}$\\Delta \\text{Log RWA}^r$ Banks", 
              "\\hspace{1.0cm}$\\Delta$ Log PD Mean",
              "\\hspace{1.0cm}$\\Delta$ Log PD Variance",
              "\\hspace{1.0cm}$\\Delta$ Log PD Skewness", 
              "\\hspace{1.0cm}$\\Delta$ Log PD Kurtosis", 
              "\\hspace{1.0cm}$\\Delta$ Log Gini",
              "$\\Delta$ Log GDP",
              #"$\\Delta$ Log Total Assets",
              #"$\\Delta$ Capital ratio",
              #"$\\Delta$ Loans / Total Assets",
              #"$\\Delta$ Net Interest Income / Oper. Rev.",
              #"$\\Delta$ Deposits / Total Assets", 
              #"$\\Delta$ Net Income / Total Assets", 
              "Log Total Assets",
              "Capital Ratio",
              "Net Income / Total Assets",
              "Loans / Total Assets",
              "Net Interest Income / Oper. Rev.",
              "Non-performing Loans / Loans",
              "Deposits / Total Assets"),
          title = "Summary statistics",
          table.layout = "n=d#-t-as-",
          notes = "\\parbox[t]{14cm}{The table shows summary statistics for banks that have introduced the IRB approach between 2007 and 2018. All variables are winsorized at 5\\%}", 
          notes.label = "",
          notes.align = "c", 
          notes.append = FALSE)
```

# Capital Savings and Business Cycle

In this section I investigate the relationship between *Capital savings* and the business cycle. Overall, capital requirements are procyclical: during economic downturns credit risk in the economy increases which raises capital requirements, under the IRB approach forcing banks to cut down lending. However, as credit conditions changes (and banks react to it) the amount of *Capital savings* also changes. The expected direction of *Capital savings* cyclicality is ambiguous. On the one hand, if the entire portfolio is equally affected, capital savings are procyclical. On the other hand, if riskier assets are more sensitive to credit shocks, then capital savings are countercyclical. The ambiguity arises because *Capital savings* can only increase, when credit conditions deteriorates, if the credit distribution widens. In the first situation, there is only a shift of the location parameters of the distribution and, consequently, a reduction of *Capital savings*.

```{r reg_portfolio, echo=FALSE}

dep.vars <- c("", ".Retail", ".Wholesale", ".Corporate", ".Sovereign", ".Banks")
aux      <- paste0("mean.PD", dep.vars)
ind.vars <- paste0("log.gdp_gr")

# Creating formulas and running the models
formula.text <- paste0(paste0("log.RWA.r", dep.vars,"_gr")," ~ ", paste0(ind.vars)," | year + bvdid | 0 | 0 ")
formulas <- lapply(formula.text, formula)

@@

```{r table_portfolio, eval=TRUE, echo=FALSE, results='asis', warning=FALSE, message=FALSE}

>>>>

---
title: "Credit Risk Distribution and Capital Savings"
author:
- name: Lucas Avezum
  affiliation: CentER and EBC, Tilburg University; l.avezum@uvt.nl; +31134664531;  Warandelaan 2, 5037 AB Tilburg, NL, The Netherlands 
date: 
header-includes:
   - \usepackage{rotating, graphicx, dcolumn, longtable, float}
output:
  pdf_document:
    #citation_package: natbib
    latex_engine: pdflatex
    template: tex-ms.tex
    fig_caption: true
  html_document:
    df_print: paged
  word_document: default
bibliography: ../../Auxiliar/mybibfile.bib
csl: ../../Auxiliar/cambridge-journal-of-economics.csl
citecolor: black
anonymous: false
fontsize: 11pt
spacing: "double"
geometry: margin= 1in
keywords: Internal ratings-based models, Risk-weighted assets dispersion, Procyclicality
JEL: G21, G28, G11
linkcolor: black
biblio-style: apsr
subtitle: ''
thanks: Preliminary draft.  
abstract: "The Basel II accord introduced a concave mapping from probabilities of default (PD) to capital requirements for banks adopting the internal ratings-based (IRB) approach. Therefore, total capital requirements depends not only on the portfolio's average PD but also on the shape of its PD distribution. Using a hand-collected dataset, I propose a method that decomposes the contribution of the PD distribution to capital requirements into an average component and a shape component. While a larger average component increases capital charges, a larger shape component generates capital savings. I document that these capital savings are important sources of risk-weight variability across banks. I further investigate the relationship of capital savings with the business cycle and find it to be countercyclical. The results suggest that a recalibration of the current IRB framework can significantly reduce the procyclicality of capital requirements."
urlcolor: black
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo      = FALSE, 
                      warning   = FALSE, 
                      error     = FALSE, 
                      tidy      = TRUE, 
                      tidy.opts = list(width.cutoff=50), 
                      collapse  = TRUE,
                      warning   = FALSE,
                      error     = FALSE,
                      message   = FALSE, 
                      comment   = "",
                      cache     = FALSE,
                      fig.pos   = 'H')
# These options are tuned for manuscript/presentation. 
# They basically run R in the background except for spitting out figures/tables
# They also cache results so that you don't have to run your whole analysis every time you fix a typo

@@

library(tidyverse)            ## Data manipulation, pipe operator
library(data.table)           ## Command to bind lists
library(ggridges)             ## Density ridges plot
library(stargazer)            ## Latex tables
library(lfe)                  ## High-dimensional fixed effects
library(gridExtra)            ## Create grid for plots and tables
library(cowplot)              ## Create grid for plots and tables
library(DescTools)            ## Command to winsorize variables
library(ggpubr)               ## More plots
#library(textablr)
#library(sandwich)
#library(lmtest)
#library(relaimpo)            ## Calculate relative importance measures

source("../../Auxiliar/Function.R") ## Capital requirements functions
source("../../Auxiliar/longtable.R") ## longtable for stargazer

@@

\begin{equation}
\label{measure1}
\text{RWA}^s=\frac{\text{RWA}^c}{\text{RWA}}.
\end{equation}

\begin{equation}
\label{measure2}
\text{RWA}^{\$}=\text{RWA}^c-\text{RWA}.
\end{equation}

\begin{equation}
\label{measure3}
\text{RW}^s=\frac{\text{RWA}^{\$}}{\text{RWA}}.
\end{equation}

The follow-up to the 2008 financial crisis offer an plausible explanation for the dynamics of *Capital savings* in time, which is illustrated in Figure \label{fig:timevariables}. Increased supervisory pressure to increase capital ratios (top plot), especially after the Basell III, led banks to deleverage and reduce risk-taking. [^1] The shift for safer assets not only reduced the average PD on banks' portfolios (middle plot) but also compressed their PD distribution (bottom plot). This distribution compression leads to a reduction in *Capital savings*. This explanation is in line with the hypothesis that *Capital savings* are countercyclical. Efforts to increase capital ratio by risk reduction are increasingly more difficult.

[^1]: Although supervisors favored raising additional capital as means to achieve the required capital ratios [@EBA2011recommendation], banks respond to capital requirements mostly by reducing RWA [@gropp2018banks].

# Capital Savings and the Business Cycle

In this section I investigate the relationship between *Capital savings* and the business cycle. I start by explaining the theoretical ambiguous relationship between *Capital savings* and the business cycle. Next I show that *Capital savings* are countercyclical, i.e., given an average portfolio PD, banks need more capital in good times than in bad times to reach the same capital ratio. Finally, I provide evidence that this relationship is not purely mechanical but depends on how banks realocate their assets through the business cycle.

## A simple theoretical framework

Overall, capital requirements are procyclical: during economic downturns credit risk in the economy increases which raises capital requirements, under the IRB approach forcing banks to cut down lending. However, as credit conditions changes (and banks react to it) the amount of *Capital savings* also changes. The expected direction of *Capital savings* cyclicality is ambiguous. On the one hand, if the entire portfolio is equally affected, capital savings are procyclical. On the other hand, if riskier assets are more sensitive to credit shocks, then capital savings are countercyclical. The ambiguity arises because *Capital savings* can only increase, when credit conditions deteriorates, if the credit distribution widens. In the first situation, there is only a shift of the location parameters of the distribution and, consequently, a reduction of *Capital savings*.

Consider now a credit risk shock. In state A, both safer and riskier assets are affected, having their PDs raised to 0.2% and 0.3%, respectively. In state B, only the riskier asset is affected, having its PD equal to 0.4%. Note that the average PD in both states are equal at 0.25%. However, relative to the initial state, the capital savings measure for state A decreases to \$0.45 while it increases to \$4.35 in state B. The mechanism of this example is simple: relative to the initial state, in state A only the location parameter of the PD distribution changes (standard deviation is equal to 0.05% in both states) while in state B also other moments change (standard deviation increases to 0.15%). Note that in both state A and B total capital requirements increase as average credit risk is higher relative to the initial state in both states. This net procyclical effect is in line with the results of @behn2016procyclical. However, due to the increased portfolio dispersion, capital savings are countercyclical in state B. Conversely, because the PD-RW formula elasticity is lower for higher PD estimates, in stata A *Capital savings* are procyclical. Therefore, capital savings are procyclical in state A. Table \ref{tab1} summarizes the example and Figure \ref{fig:pdmapcases}

## Empirical Results

In Table \ref{tab1}, I regress the change in the logarithm of *Capital savings* on the change in the logatithm of real GDP per capita. In all regressions the measure of *Capital savings* follows equation \ref{eq4}. The difference between regressions in the table is the portfolio used to calculate the *Capital savings*. The real GDP per capita used  comes from the country where the bank has its headquarter. All regressions include bank and year fixed effects and standard errors clustered at the country level.

Column 1 shows that total *Capital savings* are countercyclical as the growth rate of tota l*Capital savings* is negatively associated to the GDP growth rate. According to regression 1 an increase of 1.6 percentage points in GDP growth rate (one standard deviation) decreases total *Capital savings* growth rate by 0.7 percentage points on average. The effect is economically sizeable representing 14% of total *Capital savings* standard deviation in the sample.

The remaining columns of Table \ref{tab1} suggest that the countercyclical relationship between GDP and *Capital savings* is mostly driven by wholesale portfolios, especially corporate portfolios. The estimated coefficient in Column 2 is negative suggesting that *Capital savings* generated by retail portfolios are also countercyclical. However, this coefficient is not statistically significant. On the other hand, the coefficient of GDP growth rate on *Capital savings* from wholesale portfolios is statistically and economically significant. According to the regression 3 one standard deviation increase in GDP growth rate decreases *Capital savings* from wholesale portfolios growth rate by 0.8 percentage points on average, which is 16% of the *Capital savings* from wholesale portfolios standard deviation. Columns 3 to 6 indicate that within the wholesale portfolio most of the countercyclical variation appears to come from corporate portfolios. Only in regression 4 the coefficient of GDP growth rate is negative and statistically significant.

## Robustness to Alternative Measures

A concern to the previous analysis is the possibility of that I could be overestimating the average impact of GDP growth rate on *Capital savings*. First, Because the measure used in Table \ref{tab1} is dimentionless, i.e., it is a measure of intensity of *Capital savings*, if the portfolio of small banks, in terms of asset, are more impacted during the business cycle than the portfolio of large banks then the cofficients in Table \ref{tab1} overestimate the impact to the average bank. To account for differences in *Capital savings* size, in Table \ref{tab2} the measure of *Capital savings* follows equation \ref{eq3}, which is a measure of *Capital savings* in monetary units.

The results shown in Table \ref{tab2} reinforce the evidence that the coutercyclicality of *Capital savings* is mostly driven by the corporate portfolio.

Second, because both measures in Tables \ref{tab1} and \ref{tab2} are relative to the IRB portfolio, if the portfolio of banks with lower share of their portfolio under the IRB approach are more impacted during the business cycle than the portfolio of high IRB share banks then the cofficients in Table \ref{tab1} overestimate the impact to the average bank. The results shown in Table \ref{tab2} reinforce the evidence that the coutercyclicality of *Capital savings* is mostly driven by the corporate portfolio.

Lastly, the method that I proposed to measure the shape component of the PD distribution could be flawed. In Table \ref{tab3} I use the well-known Gini coefficient to measure the shape component of the PD distribution. Again, the results shown in Table \ref{tab3} suppport the evidence from the last three tables: *Capital savings* are countercyclical and mostly driven by the effect of GDP growth rate on *Capital savings* from corporate portfolios.

## Capital Savings and Portfolio Realocation

The results from Tables \ref{tab1} to XX raises the questions of why it appears that only corporate portfolios are impacted during the business cycle. 
I further analyse the relationship of the PD distribution moments with GDP and find that during economic expasions: average, variance and kurtosis decrease while the skewness increases. This evidence is in line with the idea that the upper-tail (riskier borrowers) of the PD distribution is disproportionately affected by the business cycle.

## Economic Impact of Capital Savings Cyclicality

# Policy Discussion: Using Capital Savings to Reduce Capital Ratio Procyclicality

# References {-}

<!-- <div id="refs"></div> -->

\newpage

# Figures {-}

@@

@@

@@

@@

```{r, fig.height = 7.9, fig.width = 6, results="hide",fig.cap= cross.dec_cap1, include=TRUE}
 plot1 <- cross.section.decomposition %>% 
  mutate(total= -PD.gain)%>%
  select(name, Country,total, int.gain, ext.gain) %>%
  gather(condition, value, contains(".gain"), factor_key=TRUE)%>%
  mutate(condition = factor(condition, levels = c("int.gain","ext.gain"),
                                       labels = c("Intensive margin", "Extensive margin")))%>%
ggplot(mapping = aes(x = reorder(name, total),
                       y = value)) +
geom_bar(stat = 'identity')+
  #geom_line(aes(y = -total,group=1))  +
  labs(x=NULL, y= NULL) +
    coord_flip() +
  theme_bw() +
  theme(panel.grid.minor = element_blank())+
  facet_wrap(~condition)

plot2 <- cross.section.decomposition %>% 
  select(name, Country, int.gain, ext.gain) %>%
  mutate_if(is.numeric,Winsorize,probs = c(0.02, 1) ,na.rm = TRUE)%>% 
  ggplot(mapping = aes(x = int.gain,
                       y = ext.gain)) +
    geom_point()+
    geom_smooth(method = "lm", color = "black")  +
  labs(x=NULL, y= NULL) +
   theme_bw() +
  theme(panel.grid.minor = element_blank())+
  scale_fill_grey()+
  stat_cor()

@@

```{r, fig.height = 7.9, fig.width = 6, results="hide",fig.cap= cross.dec_cap2, include=TRUE}
plot1 <- cross.section.decomposition %>% 
  mutate(total= -PD.gain)%>%
  select(name, Country,total, other.gain, PD.gain) %>%
  gather(condition, value, contains(".gain"), factor_key=TRUE)%>%
  mutate(condition = factor(condition, levels = c("PD.gain","other.gain"),
                                       labels = c("Capital savings", "Other components")))%>%
ggplot(mapping = aes(x = reorder(name, total),
                       y = value)) +
geom_bar(stat = 'identity')+
  labs(x=NULL, y= NULL) +
    coord_flip() +theme_bw()+
  theme(panel.grid.minor = element_blank())+
  facet_wrap(~condition)

plot2 <- cross.section.decomposition %>%
  select(name, Country, other.gain, PD.gain) %>%
  mutate_if(is.numeric,Winsorize,probs = c(0.02, 1) ,na.rm = TRUE)%>% 
  ggplot(mapping = aes(x = PD.gain,
                       y = other.gain)) +
    geom_point()+
    geom_smooth(method = "lm", color = "black")  +
  labs(x=NULL, y= NULL) +
   theme_bw() +
  theme(panel.grid.minor = element_blank())+
  scale_fill_grey()+
  stat_cor(label.y= -0.1)

@@

unrestrict <- cross.section.decomposition %>% 
  mutate(epsilon = rnorm(nrow(.),0,0.0000001),
         total.gain = total.gain+epsilon)%>% 
  lm(data = ., total.gain ~ IRB.gain + mix.gain + rollout.gain + int.gain + ext.gain + SA.gain)%>%
  relaimpo::calc.relimp(type = c("lmg"), rela = FALSE )

@@

@@

<!-- \begin{table}[!htbp] \centering  -->
<!--   \caption{Cyclicality of Capital Savings}  -->
<!--   \label{tab1}  -->
<!-- \small  -->
<!-- \begin{tabular}{@{\extracolsep{3pt}}lD{.}{.}{-3} D{.}{.}{-3} D{.}{.}{-3} D{.}{.}{-3} D{.}{.}{-3}}  -->
<!-- \hline  -->
<!-- \hline \\[-1.8ex]  -->
<!-- \\[-1.8ex] & \multicolumn{3}{c}{States}  & \multicolumn{2}{c}{$\Delta$ Initial}  \\ \cline{2-4}\cline{5-6} -->
<!-- \\[-1.8ex] & \multicolumn{1}{c}{Initial} & \multicolumn{1}{c}{A} & \multicolumn{1}{c}{B} & \multicolumn{1}{c}{A} & \multicolumn{1}{c}{B} \\  -->
<!-- \hline \\[-1.8ex]  -->
<!--  $\text{RWA}$  & 63.88 & 94.21 & 90.29 & 30.32 & 26.42 \\  -->
<!--   $\text{RWA}^c$ & 64.65 & 94.65 & 94.65 & 29.96 & 29.96  \\  -->
<!--   $\text{RWA}^s$ & 0.81 & 0.45 & 4.35 & -0.36 & 3.54  \\  -->
<!--   \hline \\[-1.8ex] -->
<!--  Avg. PD & 0.15\% & 0.25\% & 0.25\% &  &   \\  -->
<!--   S.d. PD & 0.05\% & 0.05\% & 0.23\% &  &   \\  -->
<!-- \hline \\[-1.8ex]  -->
<!-- \end{tabular}  -->
<!-- \end{table}  -->
```{r winsorized_data, fig.height = 10, fig.width = 7, results="asis",out.width = "1\\textwidth"}
## Set winsorizing limits
p_w <- c(0.05, 0.95)        
bank.data.win <- bank.data %>% 
  ungroup()%>%
  filter(!is.na(RW.s), bvdid %in% pillar3.data$bvdid) %>% 
  mutate_if(is.numeric,Winsorize,probs = p_w ,na.rm = TRUE)%>%
  group_by(bvdid)

``` 
# Tables {-}

```{r reg_cyclicality_main, echo=FALSE}

# Variables in the model 
# Dependent variables
dep.vars <- c("", ".Retail", ".Wholesale", ".Corporate", ".Sovereign", ".Banks")
dep.text <- c("$j = \\text{Total}$","$j = \\text{Retail}$", "$j = \\text{Wholesale}$", "$j = \\text{Corporate}$", "$j = \\text{Sovereign}$", "$j = \\text{Banks}$")
# Independent variables
ind.vars <- c("log.gdp_gr")
ind.text <- c("$\\Delta\\text{Log GDP}_{i,t}$")

fe.list <- list(c("Bank FE",    rep("\\multicolumn{1}{c}{Yes}", length(dep.vars))),
                c("Year FE",    rep("\\multicolumn{1}{c}{Yes}", length(dep.vars))))

# Model formulas
formula.text <- paste0(paste0("log.RWA.r", dep.vars,"_gr")," ~ ", paste0(ind.vars)," | year + bvdid | 0 | Country ")
formulas <- lapply(formula.text, formula)

```{r table_cyclicality_main, eval=TRUE, echo=FALSE, results='asis', warning=FALSE, message=FALSE}
stargazer(reg.list,
          title = "Capital Savings and the Business Cycle",
          table.layout = "n=c=d#-t-as-",
          notes = "\\parbox[t]{16cm}{The table shows estimates for the following model: \\newline \\begin{equation*}\\Delta \\text{Log RWA}_{i,t}^{s,j} = \\beta \\Delta\\text{Log GDP}_{i,t}+\\alpha_{i}+\\alpha_{t}+\\varepsilon_{i,t}. \\end{equation*} \\newline The dependent variable is the change in the logarithm of \\textit{Capital savings} for portfolio $j$. In all regressions the measure of \\textit{Capital savings} follows equation \\ref{measure1}. $\\Delta\\text{Log GDP}_{i,t}$ is the real GDP per capita growth rate. All regressions include bank and year fixed effects. Robust standard errors adjusted for clustering at the country level are reported in parentheses. *, **, and *** indicate statistical significance at the 10\\%, 5\\%, and 1\\% level respectively.}",
          dep.var.labels   = dep.text,
          covariate.labels = ind.text,
          #se               = se.list,
          add.lines        = fe.list,
          keep.stat        = c("n","rsq"),
          header           = FALSE, 
          type             = "latex",
          font.size        = "small",
          notes.label      = "",
          notes.align      = "c", 
          notes.append     = FALSE,
          label            = "tab1",
          style            = "qje",
          column.sep.width = "1pt",
          align            = TRUE)
```

\newpage

\begin{table}[!htbp] \centering  
\caption{Capital Savings and the Business Cycle: Robustness to Alternative Measures}  
\label{tab2:reg}
\footnotesize
\begin{tabular}{@{\extracolsep{1pt}}lccccc}
\multicolumn{6}{c}{\parbox[t]{16cm}{The table shows estimates for the following model: \newline \begin{equation*}\Delta \text{Y}_{i,t}^{j} = \beta \Delta\text{Log GDP}_{i,t}+\alpha_{i}+\alpha_{t}+\varepsilon_{i,t}. \end{equation*} \newline 
The dependent variable is the change in the logarithm of \textit{Capital savings} for portfolio $j$. In Panel A the measure of \textit{Capital savings} follows equation \ref{eq5}. In Panel B the measure of \textit{Capital savings} follows equation \ref{eq6}. In Panel C the Gini coefficient is used as alternative measure of the shape of the PD distribution. $\Delta\text{Log GDP}_{i,t}$ is the real GDP per capita growth rate. All regressions include bank and year fixed effects. Robust standard errors adjusted for clustering at the country level are reported in parentheses. *, **, and *** indicate statistical significance at the 10\%, 5\%, and 1\% level respectively.}} \\
[-1.8ex] \\
\hline \\
\end{tabular}
\begin{tabular}{@{\extracolsep{1pt}}cccccc}
\multicolumn{6}{c}{Panel A: $\text{Y}=\text{Log RWA}^{\$}$} \\
\end{tabular}

```{r reg_cyclicality_alt, echo=FALSE}

# Creating formulas and running the models
formula.text <- paste0(paste0("log.RWA.s", dep.vars, "_gr")," ~ ", paste0(ind.vars)," | year + bvdid | 0 | Country ")
formulas <- lapply(formula.text, formula)

@@

```{r table_cyclicality_alt, eval=TRUE, echo=FALSE, results='asis', warning=FALSE, message=FALSE}

stargazer(reg.list,
          table.layout     = "=c=d#-t-as-",
          dep.var.labels   = dep.text,
          covariate.labels = ind.text,
          #se              = se.list,
          add.lines        = fe.list,
          keep.stat        = c("n","rsq"),
          header           = FALSE, 
          type             = "latex",
          font.size        = "footnotesize",
          notes.label      = "",
          notes.align      = "c", 
          notes.append     = FALSE,
          label            = "tab2",
          style            = "qje",
          column.sep.width = "2pt",
          float            = FALSE,
          align            = TRUE)
